Matching the Time of Emergence of Transboundary Fish Stocks to Lead Time for

### 2020-03-25 Scripts/Initial_Analysis.Rmd
Analysis for estimating the proportion change

# Set paths depending on machine Beast (jepa88), "carmelia" or Hall1000
if(Sys.info()[7] == "jepa88"){
  data_path <- "Z:/DATA/DBEM/"
}
if(Sys.info()[7] == "carmelia"){
  data_path <- paste(here(),"/Temporal_Data/",sep = "")
}
if(Sys.info()[7] == "hall1000"){
  data_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Data/"
}

@@

#### The function
dbem_import <- function(taxon_key,
                        iyr,
                        fyr, 
                        ensamble_list,
                        data_type,
                        Path = NA){

  # Creat a sequence of the years to import  
  yr <- seq(iyr,
            fyr,
            by=1) #<- change this setting if you have bi/tri/etc annual data

  #### Set the Path####
  #First, set the path of the files under a loop for each year

  for(c in 1:length(ensamble_list)){

    D_Path <- paste(Path,"mpa0F1ENS",ensamble_list[c],"/", #<- drobo path for 8.5
                    taxon_key,"/",taxon_key,data_type,yr,".txt",
                    sep="")

     # total <- NULL

    #### Importing data ####
    #Now we import the data using data.table::fread and the path generated above

    for (i in 1:length(yr)){ #<-The years you want to have

      possibleError <-  tryCatch({ # <- This will solve for situations where the species is not found in the DBEM
        # Imports dataset 
        cur <- fread(D_Path[i],
                     na.strings = "NA",
                     col.names = c("index",
                                   yr[i]) #<- Calls colunmns by year
        )

      },
      error=function(e){ # Part of the "no species in DBEM
        e
      }
      )
      if(!inherits(possibleError, "error")){

        # possibleError_I <-  tryCatch({
        #Puts them together in one nice dataset 
        if(i == 1){
          total<- copy(cur) # <- copies the previouse data.table
        }else{ 
          setkey(total, # <- sets the data.table as "reference" ?setkey
                 index); setkey(cur,
                                index); 
          total <- merge(total, #<- Merges all the data in one single file.
                         cur,
                         all = TRUE) # <- This allows for the NA's

        } #Close datatable loop

      }

    } #First loop END (Path Creation)

    # Tidy version
    total <- total %>%
      gather(year,value,-index) %>%
      mutate(taxon_key = taxon_key,
             ensamble=ensamble_list[c])

    world_data <- world_grid %>%
      full_join(total,
                by = "index")
    # Messages
    if(inherits(possibleError, "error")){

      print(paste("Oops! --> No information for",taxon_key,yr[i], sep = " ")) # <- Display error message

    }else{

      print(paste("Analysis done for",taxon_key, "ensamble" ,ensamble_list[c]))
    }

    if(c == 1){
      partial_ensamble <- world_data
    }else{
      spp_ensamble <- bind_rows(partial_ensamble,world_data)
    }
  }

  return(spp_ensamble)

} #Function end

# Test function (works in Carmelia!)

x <- dbem_import(
  taxon_key <- 600004,
  iyr <- 1951,
  fyr <- 1952,
  data_type <- "Catch",
  # ensamble_list = c(102,103),
  ensamble_list =ensamble_list,
  Path <- DBEM_path
)

unique(x$ensamble)
# head(x)

# Test multiple ensambles and spp

# ensamble_list <- c(102,103)
# taxon_key_list <- c(600004,600005)

# x <- bind_rows(
#   lapply(
#     # taxon_key_list,
#     ensamble_list,
#     FUN= dbem_import,
#     taxon_key = 600004,
#     # ensamble = 102,
#     iyr = 1951,
#     fyr = 1952,
#     data_type = "Abd",
#     Path = paste(data_path,"DBEM/",sep="")
#   )
# )

# OK!

@@

EstPropChange <- function(Spp,iyr,fyr,data_type,ensamble_list,Path,Neighbours){

  # Get model data from spps
  SppDist <- dbem_import(
    taxon_key = Spp,
    iyr <- iyr,
    fyr <- fyr,
    data_type <- data_type,
    ensamble_list = ensamble_list,
    Path <- Path
  )

@@

@@

@@

  per_change_d <- full_join(Territory_T,
                            Neighbour_T, 
                            by = c("Name","taxon_key","year","ensamble")
  ) %>%
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(eez_name = Name,
           eez_neighbour =eez_name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbour = n_spp_eez.y) %>% 
    mutate(area_index = n_spp_Country/Spp_Total) %>% 
    select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
    mutate(
      time_step = ifelse(year < 2000,"historical",
                         ifelse(year >= 2045 & year <= 2064, "mid",
                                ifelse(year >= 2080, "n_end","NA")
                         )
      )
    ) %>%
    filter(time_step != "NA") %>% 
    # Time mean and sd
    ungroup() %>% 
    group_by(taxon_key,ensamble,eez_name,eez_neighbour,time_step) %>% 
    summarise_at(vars("area_index"),
                 c(temporal_mean=mean,
                   temporal_sd=sd)
    ) %>% 
    # Ensamble mean and sd
    group_by(taxon_key,eez_name,eez_neighbour,time_step) %>% 
    summarise_at(vars("temporal_mean"),
                 c(ensamble_mean=mean,
                   ensamble_sd=sd)
    ) %>% 
    mutate(ensamble_mean = ifelse(is.na(ensamble_mean),0,ensamble_mean)) %>% 
    select(-ensamble_sd) %>% 
    spread(time_step,ensamble_mean) %>% 
    mutate_at(vars(mid,n_end),
              funs(ifelse(historical == 0 & . == 0, 0, # If for no change than 0
                          ifelse(historical == 0 & . > 0, 100, # If changes from 0 to anything then 100
                                 ((.-historical)/historical)*100 # Estimate percentage
                                 # ((.-historical)/((.historical)/2))*100 # Estimate percentage
                          )
              )
              )
    )

  #### Statistical test ###

  stat_change_d <- full_join(Territory_T,
                             Neighbour_T, 
                             by = c("Name","taxon_key","year","ensamble")
  ) %>%
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(eez_name = Name,
           eez_neighbour =eez_name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbour = n_spp_eez.y) %>% 
    mutate(area_index = n_spp_Country/Spp_Total) %>% 
    select(taxon_key,year,ensamble,eez_name,eez_neighbour,area_index) %>% 
    mutate(
      time_step = ifelse(year < 2000,"historical",
                         ifelse(year >= 2045 & year <= 2064, "mid",
                                ifelse(year >= 2080, "n_end","NA")
                         )
      )
    )

  return(per_change_d)

}

@@

iyr <- 1951
fyr <- 2099
data_type <- "Catch"
ensamble_list <- seq(102,111,1)
DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
world_grid <- tibble(index = seq(1,259200,1))
Neighbours <- read_excel(paste(data_path,"Spatial/EEZ_Neighbour_List.xlsx",sep="")) %>% 
  clean_names()

@@

Spp_list <- unique(Transboundary_spp$taxon_key)[4:5]

@@

## Main Routine

```{r main_routine, eval = T, echo = F}

# Problema con 600006

####_______________________ ####
# Estimating proportion change
####_______________________ ####
r_proportion_change <- NULL
# suppressWarnings(
  r_proportion_change <- bind_rows(
    lapply(
      Spp_list, #c(600004,600005),
      EstPropChange,
      iyr <- iyr,
      fyr <- fyr,
      data_type <- "Catch",
      ensamble_list = ensamble_list,
      Path =DBEM_path,
      Neighbours = Neighbours
    )
  )
# )

write_csv(r_proportion_change,
          "r_proportion_change.csv")

>>>>

# Set paths depending on machine Beast (jepa88), "carmelia" or Hall1000
if(Sys.info()[7] == "jepa88"){
  data_path <- "Z:/DATA/DBEM/"
}
if(Sys.info()[7] == "carmelia"){
  data_path <- paste(here(),"/Temporal_Data/",sep = "")
  DBEM_path <- paste(data_path,"DBEM/",sep="")
  result_path <- paste(here(),"/Temporal_Data/Results/",sep="")
}
if(Sys.info()[7] == "hall1000"){
  data_path <- "/Volumes/DATA/JULIANO_NEYMAR/TransEmergence/Data/"
  DBEM_path <- "/Volumes/DATA/DATA/DBEM/"
  result_path <- paste(here(),"/Results/",sep="")
}

@@

dbem_import <- function(taxon_key, #Except taxon_key 
                        year, #Expects sequence of years 
                        ensamble, #Expects a number of the ensamble
                        data_type, #Expects either "Abd" or "Catch"
                        path = "NA"){

  years <- year_df %>% pull(year)

  D_Path <- paste(path,"mpa0F1ENS",ensamble,"/",taxon_key,"/",taxon_key,data_type,years,".txt",sep="")

  #### Importing data ####
  cur <- lapply(D_Path, FUN=fread, na.strings="NA")

  if(length(cur)>0){
    cur <- cur[sapply(cur, function(d) nrow(d) >= 1)] 
    colnames <- c("index", "value")
    cur <- lapply(cur, setNames, colnames)
    df <- bind_rows(cur, .id = "column_label")
    if(nrow(df)>0){
      frame_key <- tibble(column_label = seq(1,length(unlist(years)),1), 
                          "year"=years) %>% 
        mutate(column_label=as.character(column_label))
      df <- left_join(df, frame_key, by="column_label") %>% select(-column_label)
      df <- df %>% mutate(data_type=data_type,
                          taxon_key = taxon_key,
                          ensamble = ensamble
      )
    }
  } else {
    df <- tibble()
  }
    return(df)
} #Function end

# x <- dbem_import(
#   taxon_key <- 600004,
#   Year <- year_df,
#   data_type <- "Catch",
#   ensamble = 102,
#   # ensamble_list =ensamble,
#   path <- DBEM_path
# )

# attach(x)
# head(x)
# unique(year)
# unique(ensamble_list)
# detach(x)

@@

EstPropChange <- function(taxon_key,
                          year,
                          data_type,
                          ensamble_list, # expects a list of ensambles to averrage by
                          path,
                          neighbours){

  # Loads species data for all ensambles
  spp_dist <- bind_rows(
    lapply(
      ensamble_list,
      dbem_import,
      taxon_key = taxon_key,
      year = year,
      data_type = data_type,
      path = path
    )
  )

  # head(spp_dist)

@@

  suppressWarnings(
    trans_EEZ <- Transboundary_spp %>% 
      filter(taxon_key %in% spp_dist$taxon_key) %>% 
      left_join(Neighbours,
                by = c("eez_name","eez_neighbour")
      )
  )

  # List of indexes per EEZ
  index_id <- trans_EEZ %>% 
    group_by(eez_name,index) %>% 
    summarise(n=n()) %>% 
    select(-n)

  # --------------- #
  # Warning messages:
  # 1: Column `eez_name` joining factor and character vector, coercing into character vector 
  # 2: Column `eez_neighbour` joining factor and character vector, coercing into character vector 
  # --------------- #

  #____________ Selecting only the transboundary nature of the species _________ #
  trans_spp <- spp_dist %>%
    filter(index %in% trans_EEZ$index) %>% 
    left_join(index_id,
              by = "index")

@@

  #Step 1.  Get EEZ id and Neighbour
  Neighbours_List <- Neighbours %>% 
    group_by(eez_name,eez_neighbour) %>% 
    summarise(n=n()) %>% 
    ungroup() %>% 
    select(-n) %>% 
    semi_join(trans_EEZ,
              by = c("eez_name","eez_neighbour")
    )

  # Step 2. Determines the amount of grids present in each country
  spp_grid <- trans_spp %>% 
    group_by(taxon_key,
             eez_name,
             year,
             ensamble) %>% 
    summarise(n_spp_eez = length(unique(index))) %>% 
    left_join(Neighbours_List,
              by = "eez_name")

@@

  # Split dataframes to merge latter
  Territory_T <- spp_grid %>% 
    ungroup() %>% 
    select(
      taxon_key,
      Name=eez_name,
      n_spp_eez,
      year,
      ensamble
    )

  Neighbour_T <- spp_grid %>% 
    ungroup() %>% 
    select(
      taxon_key,
      n_spp_eez,
      Name=eez_neighbour,
      eez_name,
      year,
      ensamble
    )

  per_change_d <- full_join(Territory_T,
                            Neighbour_T, 
                            by = c("Name","taxon_key","year","ensamble")
  ) %>%
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per Neighbours
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(eez_name = Name,
           eez_neighbour =eez_name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbour = n_spp_eez.y) %>% 
    mutate(area_index = n_spp_Country/Spp_Total) %>% 
    # select(taxon_key,year,ensamble_list,eez_name,eez_neighbour,area_index) %>% 
    mutate(
      time_step = ifelse(year < 2000,"historical",
                         ifelse(year >= 2045 & year <= 2064, "mid",
                                ifelse(year >= 2080, "n_end","NA")
                         )
      )
    ) %>%
    filter(time_step != "NA") %>% 
    # Time mean and sd
    ungroup() %>% 
    group_by(taxon_key,ensamble,eez_name,eez_neighbour,time_step) %>% 
    summarise_at(vars("area_index"),
                 c(temporal_mean=mean,
                   temporal_sd=sd)
    ) %>% 
    # Ensamble mean and sd
    group_by(taxon_key,eez_name,eez_neighbour,time_step) %>%
    summarise_at(vars("temporal_mean"),
                 c(ensamble_mean=mean,
                   ensamble_sd=sd),
                 na.rm=T
    )

  # --------------- #
  # End function
  # --------------- #

  if(nrow(per_change_d) > 0){

  # Save Result
  name <- paste("proportion_",taxon_key,".csv",sep="")
  path_name <- paste(result_path,"Proportion/",name, sep="")

  write_csv(per_change_d,
            path_name)

  return(print(paste("analysis done",taxon_key)))

  }else{
    print(paste("no result for",taxon_key))
  }

  # Or return a df
  # return(per_change_d)
}

@@

# Data needed for estimating Proportion change

year_df <- tibble(year =  c(seq(1995,2014,1), seq(2041,2060,1), seq(2080,2099,1)),
                  time_period = c(rep("Today", 20), rep("Mid Century", 20), rep("End of Century", 20))
                  )

data_type <- "Catch"
ensamble_list <- seq(102,103,1)
world_grid <- tibble(index = seq(1,259200,1))
Neighbours <- read_excel(paste(data_path,"Spatial/EEZ_Neighbour_List.xlsx",sep="")) %>% 
  clean_names()

@@

spp_list <- unique(Transboundary_spp$taxon_key)[1:3]

@@

## Proportion Routine

```{r Proportion_routine, eval = T, echo = F}
####_______________________ ####
# Estimating proportion change
####_______________________ ####

lapply(
  spp_list, #c(600004,600005),
  EstPropChange,
  year <- year_df,
  data_type <- "Catch",
  ensamble_list = ensamble_list,
  path = DBEM_path,
  neighbours = Neighbours
)
